---
title: "The pulsatile user's guide"
author: 
  - name: Matt Mulvahill 
    email: matthew.mulvahill@ucdenver.edu
    affiliation: CU Anschutz
package: pulsatile
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The pulsatile user's guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



# Introduction



```{r options, echo=FALSE}
library(knitr)
knitr::opts_chunk$set(collapse = TRUE,
                      echo     = TRUE,
                      message  = TRUE,
                      warning  = TRUE,
                      eval     = TRUE)
```



`pulsatile` is an R package for Bayesian modeling of pulsatile hormone time
series data via birth-death MCMC. BDMCMC is an unsupervised statistical learning
algorithm for estimating mixture models with unknown numbers of components. 

The data consists of hormone concentration measurements from blood samples
typically taken every 6-10 minutes for 6-24 hours.  With this specific analysis
we are attempting to estimate the number and location of hormone pulses in the
sample, as well as individual and sample-mean pulse features like quantity of
hormone released (mass), duration of release (width), elimination rate
(half-life), and baseline hormone levels.

This package currently models one subjects time-series.  We are currently
refactoring and developing the population models and two-hormone (driver and
response hormones) models.

The development version is available on
[GitHub](https://github.com/BayesPulse/pulsatile.git).  


#### Example analysis

```{r install-pulsatile}
# Install/load pulsatile with devtools
if (!require(devtools))  { install.packages("devtools")           }
if (!require(pulsatile)) { install_github("BayesPulse/pulsatile") }
```

```{r load-otherpkgs}
# Install/load other packages
if (!require(tidyr))    { install.packages("tidyr")    }
if (!require(dplyr))    { install.packages("dplyr")    }
if (!require(ggplot2))  { install.packages("ggplot2")  }
if (!require(ggthemes)) { install.packages("ggthemes") }
if (!require(coda))     { install.packages("coda") }

# Set options
theme_set(theme_fivethirtyeight())
options(scipen = 999)
```



```{r generate-sim}
# Generate (reproducible) simulated pulsatile hormone time-series
set.seed(2018-01-23)
this_pulse <- simulate_pulse()

# View/inspect simulated data
str(this_pulse)
plot(this_pulse)
```

```{r fit}
# Create model specification object (for priors, starting values, and proposal
# variances)
model_spec <- pulse_spec(location_prior_type = "order-statistic")

# Fit model (takes roughly 3-10 minutes) 
#   Note: fit_pulse accepts a plain dataset with the time and concentration
#         variable names specified, or an object created by simulate_pulse()
fit_test   <- fit_pulse(.data = this_pulse, 
                        iters = 25000, 
                        thin = 50,
                        spec = model_spec, 
                        verbose = FALSE)
```

```{r inspect}
# Inspect posterior chains
str(fit_test, max.level = 1)
common_chain(fit_test)
pulse_chain(fit_test)
```

```{r diagnostics}
# Review traceplot and posterior distributions
bp_trace(fit_test)
bp_posteriors(fit_test)
bp_location_posterior(fit_test)
```

```{r summarystats}
# Calculate mean and 80% HPD credible interval
rbind(
      common_chain(fit_test) %>% 
        summarise_at(vars(num_pulses:sd_widths), mean),
      common_chain(fit_test) %>% 
        select(-iteration) %>% 
        as.mcmc %>% 
        HPDinterval(., prob = 0.8) %>% t
      ) %>%
mutate(stat = c("mean", "lower", "upper")) %>%
gather(key = parameter, value = value, -stat) %>%
spread(key = stat, value = value)


```


# Session Info

```{r sessionInfo}
sessionInfo()
warnings()
```
